---
title: "assigh01-roulette"
author: " Wenqi Lyu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    code_folding: hide
    toc: yes
    number_sections: true
    toc_depth: 3
    toc_float: true
---

incomplete. Please follow assignment instructions:

You should explain how you used computer simulation to calculate the
average earnings of a gambler that uses this strategy. As part of
the explanation, provide a figure (or a series of figures) that show
how the gamblers earnings (or losses) evolve over a series of wagers
at the roulette wheel. (The x-axis will be the wager number (or play
number), the y-axis will be earnings.) The code below provides all
the functions you’ll need to calculate average earnings.

Show your audience how changing a parameter of the simulation (see
table below) does or does not have an impact on average earnings. A
figure would be helpful.

See the stopping rule below. Explain to your audience how you used
computer simulation to estimate the average number of plays before
stopping. The code below will need to be modified to calculate this
quantity.

Be sure to explain the limitations of the simulation; identify
simplifications or other sources of uncertainty.

```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

# Introduction
There are several strategies for playing roulette. (Strategies is in italics because the house always wins, in the long run.) Consider one such strategy:“Martingale” strategy.

## Background
Here is an image could help you understand"Martingale":

![Caption: Image of Martingale.](/Users/sevenlv/Projects/probability-and-inference-portfolio-wenqi-lyu/01-roulette/martingale-strategy.svg){#id .class width=50% height=50%}

# Methods
```{r}
library(magrittr)
#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 

for(i in c(1:3)){
  one_person_play <- one_series(B = 200, W = 300, L = 1000, M = 100)
  
  plot(x=c(1:length(one_person_play)), y=one_person_play, type="l", xlab = "Play Number", ylab = "Money in Pocket")
}


```
## Simulation

```{r}

# Simulation
walk_out_money <- rep(NA, 100)
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_last
}

# Walk out money distribution
hist(walk_out_money, breaks = 100)

# Estimated probability of walking out with extra cash
#mean(walk_out_money > 200)

# Estimated earnings
#mean(walk_out_money - 200)
```

The Estimate probability of walking out with extra cash is `r mean(walk_out_money > 200)`. The Estimate earnings would be on average `mean(walk_out_money - 200)`. Our estimated earnings are negative and we walk out with more money less than 1/2 of the time on average. We might lose some money on this game, we could change another strategy which is more profitable.

## Changing the the maximum number of plays 

L = 1000 changed to L=10, I hope to see that more people walk out with more than $0
```{r}

# Simulation
walk_out_money <- rep(NA, 100)
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 10, M = 100) %>% get_last
}

# Walk out money distribution
hist(walk_out_money, breaks = 100)

# Estimated probability of walking out with extra cash
#mean(walk_out_money > 200)

# Estimated earnings
#mean(walk_out_money - 200)
```
The Estimate probability of walking out with extra cash is `r mean(walk_out_money > 200)`. The Estimate earnings would be on average `r mean(walk_out_money - 200)`. more people will walk out with extra money than the simulation mentioned before.


### optional plot
```{r}
library(magrittr)
max_num_plays = seq(10,1000, by=10)
avg.earn= NA

for(i in c(1:length(max_num_plays))){
  #Simulation
  walk_out_money <- rep(NA,100)
  for (j in seq_along(walk_out_money)){
    walk_out_money[j] <- one_series(B = 200, W= 300, L = max_num_plays[i], M = 100) %>% get_last
  }
  #Estimated earnings
  avg.earn[i] = mean(walk_out_money - 200)
}

plot(x=max_num_plays, y=avg.earn, type="l")
```


# Stopping Rule

```{r}
# Simulation
num.plays <- rep(NA,100)
for(j in seq_along(num.plays)){
  num.plays[j] <- one_series(B = 200, W =300, L = 1000, M = 100) %>% length()
}


#Walk out money distribution

hist(num.plays, breaks = 10)
#mean(num.plays)
```
the mean number of plays is `r mean(num.plays)` , because this is so close to L=10 we know almost everyone is stopping because they reach the max of plays.

# Conclusion
Playing roulette in the (B = 200, W =300, L = 1000, M = 100 scenario is high risk to lose money. But when the L decreases, more people will walk out with extra money than the simulation mentioned before.

# Limitation
the conclusion might be biased if we do not simulate a large amount of people participant in the roulette.